\documentclass[master,english]{include/tfgitic}

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{array}

%% ADDED FOR ITSC IMPORT
\usepackage{import}
\usepackage[capitalise]{cleveref}
\usepackage[noend,noline,plainruled]{algorithm2e}
\usepackage{annotate-equations}
\usepackage{circledsteps}
\usepackage{mathtools}
\usepackage{transparent}

\DontPrintSemicolon
\SetAlgoCaptionSeparator{.}
\SetAlCapNameFnt{\footnotesize}
\SetAlCapFnt{\mdseries \footnotesize}
\SetAlCapSkip{0.5em}
\SetAlgoInsideSkip{medskip}
\SetKwSty{textsc}
\SetKwInput{KwData}{\mdseries\textsc{data}}
\SetKwInput{KwResult}{\mdseries\textsc{result}}
\crefname{algocf}{Alg.}{Algs.}
\Crefname{algocf}{Alg.}{Algs.}

\tikzset{annotate equations/arrow/.style={-, shorten <=2pt}}
\DeclareMathOperator*{\minimize}{minimize}
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
%% END ADDED FOR ITSC IMPORT

\addbibresource{misc/thesis.bib}

\title{Estimating games \\of ordered preference}
\subtitle{Scalable preference-order estimation \\for Interactive Autonomous Systems}

\author[m]{Eric Roy-Almonacid}
\advisor{Georgios Bakirtzis}
\counselor{Aleix Llusà Serra}
\company{Institut Polytechnique de Paris}
\school{Universitat Politecnica de Catalunya}
\degree{Master's degree in Machine Learning and Cybersecurity for Internet Connected Systems}
\location{Manresa (Barcelona), Spain}
\ects{18}
\topics{multiobjective optimization, game-theoretic planning, autonomous driving}

\begin{acknowledgments}
I would like to thank my advisor, Georgios Bakirtzis, for introducing me into
the research world and allowing me to attend a conference and a summer school.
I also want to thank Pau de las Heras Molins for allowing me to work together
in this exciting (but also intense) journey.
This thesis wouldn't have gone that far without you both.

Moreover, this project wouldn't have seen the light if it wasn't for the
wonderful team that helped Pau and I submit our first paper, allowed us to
participate in their other projects, and gave feedback on our future works:
David Fridovich-Keil, Lasse Peters, Dong Ho Lee, and Georgios Bakirtzis.

I wish to thank too the board of the UPC-MERIT project and EPSEM's personnel to
partially fund my attendance to the conference and summer school. 
This work is partially supported by the E.U. Digital Europe Programme under grant
DIGITAL-101083531-MERIT.

Finally, I want to thank my family and friends for supporting my career and for
understanding my lack of social life these past months.

% RAL
% Autonomous vehicles must balance ranked objectives, such as minimizing
% travel time, ensuring safety, and coordinating with traffic.
% Games of ordered preference effectively model these interactions, but
% require other agents to share their ranked objectives in order to efficiently calculating optimal trajectories --- which is incompatible with scenarios with human agents. While prior work efficiently estimates
% some of the other agents' game parameters, like goal position, there are no formal methods of inferring preference relations from some (partial) observations. In this paper we close this gap by doing X. We demonstrate through simulated traffic scenarios and small-scale robots that thinking that other agents have preference relations enable for better agent predictions.
\end{acknowledgments}

\begin{resum}
Els sistemes autònoms sovint han de satisfer diversos objectius que no es
poden reduir de manera natural a un únic cost escalar. Això és especialment
evident en la conducció autònoma, on la seguretat, el compliment de les
normes, el manteniment del carril i el progrés cap al destí s'han de
prioritzar de manera diferent segons la situació. Aquesta tesi estudia
aquest problema mitjançant els jocs de preferència ordenada, un marc de
teoria de jocs en què cada agent ordena lexicogràficament els
seus objectius.

La tesi se centra en tres aspectes necessaris per fer aquest marc més útil
a la pràctica. En primer lloc, es desenvolupa \emph{lexicographic IBR over
time}, un mètode aproximat de resolució en horitzó lliscant que aprofita
la continuïtat temporal per reduir el cost computacional de resoldre jocs
de preferència ordenada, mantenint l'efecte qualitatiu de l'ordre de
preferències sobre el comportament. En segon lloc, es formula el problema
invers d'estimar l'ordre de preferències d'altres agents a partir
d'observacions, i posa de manifest la limitació de l'assumpció
d'informació completa habitual en els jocs de trajectòries. Finalment,
presenta Duckietown com a plataforma experimental de baix cost per
avançar cap a una validació física dels mètodes proposats sota soroll de
mesura, errors d'actuació i altres imperfeccions del món real.

En conjunt, els resultats mostren que els jocs de preferència ordenada
constitueixen un marc prometedor per modelar la presa de decisions
interactiva amb objectius jerarquitzats, i que les aproximacions en
horitzó lliscant poden fer-lo considerablement més tractable. Alhora, la
tesi identifica l'estimació de preferències i una validació \emph{sim-to-real}
més àmplia com a línies essencials de treball futur.
\end{resum}

\begin{abstract}
Autonomous systems often operate under multiple objectives that cannot be
naturally reduced to a single scalar cost. This is especially true in
autonomous driving, where safety, rule compliance, lane keeping, and
progress towards a destination must be prioritized differently depending
on the situation. This thesis studies that problem through games of
ordered preference, a game-theoretic framework in which each agent ranks
its objectives lexicographically.

The thesis focuses on three aspects required for the practical use of this
framework. First, it develops \emph{lexicographic IBR over time}, an
approximate receding-horizon solution method that exploits temporal
continuity to reduce the computational burden of solving games of ordered
preference while preserving the qualitative influence of preference order
on behavior. Results in simulated traffic scenarios show substantial
reductions in compilation and solution times relative to the baseline
formulation, especially as the number of preference levels increases.
Second, it formulates the inverse problem of estimating the preference
order of other agents from observations, highlighting the need to move
beyond the complete-information assumption common in trajectory games.
Third, it introduces Duckietown as a low-cost robotic platform for moving
towards experimental validation under sensing noise, actuation errors, and
other real-world imperfections.

Taken together, the thesis shows that games of ordered preference provide
a promising framework for modelling interactive decision making with
prioritized objectives, and that receding-horizon approximations can make
their use considerably more tractable. At the same time, it identifies
preference-order estimation and broader sim-to-real validation as
essential directions for future work.
\end{abstract}

\begin{document}

\include{chapters/01_introduction}
\include{chapters/02_background}
\include{chapters/03_goop}
\include{chapters/04_lex_ibr_ot}
\include{chapters/05_estimation}
\include{chapters/06_duckietown}
\include{chapters/07_conclusions}

\chapter*{Endnotes}

Chapters \ref{ch:goop} and \ref{ch:lex-ibr-ot} are edited from a publication co-first-authored by Pau de las Heras Molins and
myself, with Dong Ho Lee, Lasse Peters, David Fridovich-Keil, and Georgios Bakirtzis as
coauthors, which was accepted and presented at the \href{https://ieee-itsc.org/2025/}
{2025 IEEE International Conference on Intelligent Transportation Systems} (ITSC 2025) \cite{heras:2025}.

\printbibliography

\end{document}
