\chapter{Background}

In this chapter we review the prior work related to this thesis.
We have been primarily motivated
to study games of ordered preference
by Zanardi et al. \cite{zanardi:2022} and Lee et al. \cite{lee:2025}.
We explicitly build on this research
to compute for a receding horizon,
rather than a finite-time horizon,
to improve the scalability of these games.
We have additionally found insights in receding horizon games
\cite{hall:2022,hall:2024}, although the formulation is not directly applicable
to the problem of preference ordering.

Prior work, closely related to the formulation
of games of ordered preference in this thesis,
is on \emph{prioritized metrics} (\cref{sec:prioritized-metrics}) 
and \emph{game-theoretic planning} (\cref{sec:game-theoretic-planning}).
In parallel, we also review prior work on \emph{game state estimation} (\cref{sec:game-state-estimation}),
and \emph{sim-to-real} (\cref{sec:sim-to-real}).

\section{Prioritized metrics}
\label{sec:prioritized-metrics}
Posetal games \cite{zanardi:2022} formalize games
with prioritized metrics
through partial orders over system designs \cite{censi:2019}, using symbolic structures to encode constraint and preference hierarchies \cite{wongpiromsarn:2021}.
While posetal games focus
on discrete decision spaces, we adopt lexicographic optimization \cite{penlington:2024},
that extends prioritized metrics to continuous action spaces \cite{lee:2025}.
However, the lexicographic nested problem structure introduces computational challenges: lower-priority objectives are optimized only after higher-priority ones are satisfied, creating a hierarchy of interdependent subproblems. To mitigate this computational intractability, prior work transcribes the lexicographic hierarchy into mathematical programs with complementarity constraints \cite{nurkanovic:2024}, with the addition of relaxation schemes \cite{lee:2025}.
Yet, coupled constraints cause scalability to remain limited.
By framing the problem
within a receding horizon setting, we exploit the temporal decoupling inherent
to IBR algorithms, which are known to be computational tractable.

\section{Game-theoretic planning}
\label{sec:game-theoretic-planning}
Game-theoretic traffic
management is needed to analyze the interactive nature of preferences between
autonomous agents \cite{qin:2024}. Efficient solution strategies exist
for game-theoretic planning \cite{laine:2021,williams:2023,mehr:2023,zhu:2023}, but solving games that involve nested optimization problems typically requires approximate solutions, similar to those used in solving trajectory games with continuous action spaces \cite{jond:2022,peters:2022,liu:2023}.
Moreover, predictions of other agents' intentions combined with planning
increase the accuracy of approximate solutions \cite{li:2022,zhang:2023}. Receding
horizon in game-theoretic planning has been studied in the classical formulation
\cite{veer:2023} and in variants---where objects are divided into immediate
and mid-term metrics \cite{fisac:2019}. These approaches do not consider the
problem of preferences in agent interactions.

\section{Game state estimation}
\label{sec:game-state-estimation}
Perception has received a lot of attention since it enables the usage of game-theoretical methods in real scenarios \cite{schwarting:2018}.
Game theory helps building a model from partial observations \cite{liniger:2020} and approximates optimal trajectories that comply with safety constraints \cite{fridovichkeil:2020}.
Goal positions can be predicted using minimum likelihood estimators \cite{liu:2023}, or by solving inverse games \cite{li:2023}, although variational auto-encoders can be used to train models that perform better overall \cite{liu:2024}. Receding horizon is commonly employed in dynamic games to take advantage of
previous observations over time \cite{peters:2023}.
Given the difficulty of predicting other's behavior, contingency planning has proven useful in many situations \cite{peters:2024}.
The inverse game problem is also a common approach for Markov gamesn reinforcement learning, and Stackelberg games \cite{chen:2023,lian:2024,ward:2025}.
In this work we use inverse games to infer the preference order.

\section{Sim-to-real}
\label{sec:sim-to-real}
The Duckietown project proposes a common environment to learn and perform experiments on autonomous vehicles \cite{duckietown:2017}. Their low cost makes it accessible to perform real tests, thus reducing the sim-to-real gap. Many works use camera and odometry input to follow the road and detect immediate obstacles, turns and intersections \cite{castro:2019}.
External inputs such as camera tracking complement the robot's sensors and allow for more complex behaviors, such or obstacle
avoidance \cite{candela:2022}.
Duckiebots are also used outside the standardized city model and are able to navigate through sidewalks or other real-world situations using reinforcement learning techniques \cite{kahn:2021}.
In this work we make robots interact with each other in intersection scenarios.
Apart from Duckietown, the CARLA simulator is vastly used to test complex environments \cite{carla:2017}, and projects like f1/10 \cite{f110:2019} also provide cyber-physical systems to perform experiments.

\section{Terminology}

The reader is referred to \cite{fridovich-keil:2024} for a smooth introduction
into game theory. In this section, we will define the basic terminology used in
this field.

\smallskip \noindent
\textbf{Agents or players} \quad
Decision-making entities that interact in a shared environment, where each agent's actions influence others.

\smallskip \noindent
\textbf{Objective or cost function} \quad
A function evaluating an agent's behavior over trajectories.
Minimizing the outcome of these functions yields better trajectories for the agent.
In this work, objectives are ordered and compared lexicographically.

\smallskip \noindent
\textbf{Trajectory} \quad
A sequence of states and actions over time generated by system dynamics and control inputs.
In non-trajectory game theoretical models, it is also called strategy.

\smallskip \noindent
\textbf{Generalized Nash equilibrium (GNE)} \quad
A set of strategies where no agent can improve its objective unilaterally, under shared and coupled constraints.

\smallskip \noindent
\textbf{Preference relations} \quad
A binary relation over outcomes encoding how agents rank them. Here, preferences are ordered and induce lexicographic comparisons.

\smallskip \noindent
\textbf{Inverse game} \quad
A method to infer agents' objectives or preferences from observed behavior in a game-theoretic setting.

\smallskip \noindent
\textbf{Iterated best response (IBR)} \quad
An algorithm where agents sequentially optimize their strategy given others' fixed strategies until convergence.

\smallskip \noindent
\textbf{Coupled constraints} \quad
Constraints that depend on multiple agents' decisions, linking their feasible strategy sets.

\smallskip \noindent
\textbf{Dynamic game} \quad
A game where agents interact over time through system dynamics and evolving states.
