\chapter{Conclusions and future work}
\label{ch:conclusions}

This chapter summarizes the main findings of the thesis and outlines
promising directions for continuing this line of research.

\section{Conclusions}

This thesis studied autonomous decision making in interactive
environments through the framework of games of ordered preference. The
motivation is that many autonomous systems, and especially autonomous
vehicles, must satisfy multiple objectives whose relative importance
cannot be represented satisfactorily by a single scalarized cost.
Preference relations provide a suitable formalism for capturing such
prioritized objectives, but their practical value depends on whether
they can be solved efficiently, inferred from observations, and
validated beyond purely idealized simulation settings.

\smallskip \noindent
\textsc{computational tractability} \quad
The principal contribution of the thesis lies in the computational
tractability of games of ordered preference. In
\cref{ch:lex-ibr-ot}, we introduced lexicographic IBR over time as an
approximate solution method for receding-horizon games of ordered
preference. The results show that the method preserves the qualitative
effect of different preference relations on agent behavior while
substantially reducing compilation and solution times with respect to the
baseline formulation. This advantage becomes particularly significant as
the number of preference levels increases, which is precisely the regime
in which the standard formulation becomes least practical. The results of
this chapter therefore support the conclusion that approximate solutions
can be obtained efficiently enough to strengthen the practical viability
of preference-aware planning in interactive traffic scenarios.

\smallskip \noindent
\textsc{preference order estimation} \quad
The second contribution of the thesis concerns the inverse problem of
games of ordered preference. Although this part remains less developed
than the computational contribution, it goes beyond identifying a missing
problem formulation. The experiments in \cref{ch:estimation} show that
exact offline recovery is difficult under short observation windows,
mainly because lower-priority objectives may never become active strongly
enough to reveal the underlying order. In that sense, the main bottleneck
of the offline study is not merely the size of the search space, but the
limited identifying signal available in the observed interaction. At the
same time, the chapter also shows a more positive result: when the true
profile belongs to a small candidate set around a good nominal prior, the
online updater can identify the correct joint non-ego order from the
first update onward, even in the tighter city scenario. This clarifies
that efficient solution methods alone are not sufficient for practical
deployment, but also that preference-order inference can still be useful
when it is posed as local discrimination rather than unconstrained
discovery.

\smallskip \noindent
\textsc{experimental validation} \quad
The third contribution of the thesis is the inclusion of experimental
validation of the receding-horizon planning stack on the Duckietown
platform. The work presented in
\cref{ch:duckietown} is intentionally modest in scope, but it already
makes the sim-to-real bottlenecks more concrete.
The experiments show that the framework can be executed in a near-real-time
physical loop. In the reported runtime logs, all but one of the solves
finish within one sampling period and all recorded solves finish before the
next replanning deadline imposed by the turn length, although the
slowest replans still leave limited slack.
% At the same time, the reported physical study covers only one interaction
% setting, so it should be interpreted as a proof of executability rather
% than as a broad comparative campaign.
The experiments also reveal a systematic gap between planned acceleration-domain
trajectories and executed robot motion, because the DB21J platform must
realize those commands through its native \((v,\omega)\) interface.
This is relevant because the significance of preference-aware planning
does not depend only on formal models and simulation results, but also on
how those methods behave under sensing noise, actuation errors,
synchronization issues, and partial observability.
The Duckietown platform should therefore be understood not only as an
enabling step, but also as an initial empirical demonstration of where
the current implementation already works and where it still needs tighter
execution.

Taken together, the results of the thesis indicate that games of ordered
preference constitute a promising framework for representing interactive
decision making with prioritized objectives, and that their use can be
made considerably more practical through receding-horizon approximations.
At the same time, the thesis also shows that further progress in
preference estimation and experimental validation is needed if the
framework is to become fully useful in realistic autonomous-driving
settings. In particular, the estimation results suggest that practical
success depends both on the richness of the observation window and on the
quality of the candidate set against which alternative preference orders
are tested.

\section{Future work}

Future work follows naturally from the three main axes of the thesis.
The most immediate opportunities lie in improving the tractability of
games of ordered preference further, broadening the methodological basis
for preference estimation, and strengthening the experimental validation
of the proposed framework.

\smallskip \noindent
\textsc{computational tractability} \quad
While this thesis achieved a considerable efficiency improvement by
introducing lexicographic IBR over time, the obtained trajectories
remain approximations of the optimal solution. Future work should
therefore look for further accelerations that preserve solution
quality.
In particular, the tractable
reformulation proposed by Lee et al. \cite{lee:2026} replaces the
exponentially growing complete KKT system with a reduced system whose
size grows polynomially with the number of players and preference levels, all
while obtaining optimal solutions.
These results suggest that structural reformulations and temporal
approximation schemes may be combined to obtain methods that are both
efficient and accurate, although doing so will still require care in
more general nonlinear settings.

\smallskip \noindent
\textsc{preference order estimation} \quad
Future work should also continue the development of inverse games of
ordered preference. This includes improving how candidate preference
orders are generated, extending the online method beyond local
discrimination around a nominal prior, and studying how longer or more
informative observation windows affect identifiability. It also includes
comparing alternative inference strategies and analyzing how partial
observability, model mismatch, and limited sensing affect the recovery of
other agents' preference orders. Since this part of the thesis remains at
an initial stage, it offers considerable room for methodological
improvement.

\smallskip \noindent
\textsc{experimental validation} \quad
Another important direction is to make the experimental evaluation more
robust and substantially broader. This involves conducting a larger
number of experiments, considering richer interaction scenarios,
incorporating additional sensing modalities, and combining the proposed
methods with other existing works in perception, estimation, and
planning. It also involves tightening the execution stack itself, in
particular by reducing solver latency and by improving the map from
planned accelerations to the robot's low-level control interface.
Extending the use of the Duckietown platform in this way would provide a
more complete assessment of how games of ordered preference behave under
realistic sim-to-real conditions.
